HO unification from object language to meta language

### 2024-03-20 paper/main.tex
some comments on 6

On the other hand, the \Ue is not designed to understand how the $\beta$-redexes
work in the onject language. Therefore, even if we know that $F$ is assigned
to $\lambda x.x$, \Ue is not able to unify $F a$ with $a$. On the other hand,
the problem $F a = G$ is solvable by \Ue, but the final result is that $G$ is
assigned to $(\lambda x.x) a$ which breaks the invariant saying that the 
substitution of the meta language always contain only terms in normal form.

@@

This link carries two terms, the former representing the variable $h$ for the
new created hole and the latter containing the subterm $t$. As for the \linketa,
we will call $h$ and $t$ respectively the left $lhs$ and the right $rhs$ hand
side of the \linkbeta.

At compile time, a subterm is \maybebeta, if it has the shape \elpiIn{fapp[fuva
N | L]} and \elpiIn{distinct L} does not hold. In that case, \elpiIn{L} is split
in two sublist \elpiIn{PF} and \elpiIn{NPF} such that former is the longest
prefix of \elpiIn{L} such that \elpiIn{distinct PF} holds. \elpiIn{NPF} is the
list such that \elpiIn{append PF NPF L}. The $lhs$ is set to a new variable
named $M$ with \elpiIn{PF} in scope whereas the $rhs$ is given by the term
\elpiIn{app[uva N' PF | NPF]} where the \Ho variable identified by \elpiIn{N'}
is mapped to the \Fo variable named \elpiIn{N}.

After its creation, a \linkbeta remain suspended until the head of $rhs$ is
instantiated by the oracle (see \cref{prop:simulation}). In this case, $rhs$ is
$\beta$-reduced to a new term, say $t'$. $t'$ is either a term in \llambda, in
which case $t'$ is unified with the $lhs$, otherwise, the link remain suspended
and no progress is performed. Another way to wake a \linkbeta up is when the
$lhs$ is a term \elpiIn{T} and $rhs$ has the shape \elpiIn{app[uva N PF | NPF]}
and some of the arguments in the \elpiIn{NPF} list become names. This is
possible after the resolution of other links. In this case, the list \elpiIn{L}
obtained by the concatenation between \elpiIn{PF} and \elpiIn{NPF} is split
again in to lists \elpiIn{PF'} and \elpiIn{NPF'}. If \elpiIn{PF} is not that
same as \elpiIn{PF'}, then we can 1) remove the current \linkbeta, 2) create a
new \linkbeta between \elpiIn{T} and \elpiIn{app[uva N' PF' | NPF']} and 3)
create a new \linketa between the variables \elpiIn{N} and \elpiIn{N'}.

An example justifying this last link manipulation is given by the following
unification problem:

@@

We can now represent the \hrun execution from this configuration which will, at
first, dereference all the links, and then try to solve them. The only link
being modified is the second one, which is set to \textIn{X3 x =~$\beta$~= X2
x a}. The $rhs$ of the link has now a variable which is partially in the PF,
we can therefore remove the original \linkbeta and replace it with the following 
couple on links:

>>>>

On the other hand, the \Ue is not designed to understand how the $\beta$-redexes
work in the object language. Therefore, even if we know that $F$ is assigned
to $\lambda x.x$, \Ue is not able to unify $F a$ with $a$. On the other hand,
the problem $F a = G$ is solvable by \Ue, but the final result is that $G$ is
assigned to $(\lambda x.x) a$ which breaks the invariant saying that the 
substitution of the meta language does not generate terms outide \wellb{} (Property \ref{prop:nf}).

@@

This link carries two terms, the former representing the variable $h$ for the
new created hole and the latter containing the subterm $t$. As for the \linketa,
we will call $h$ and $t$ respectively the left hand side (LHS)
and the right hand side (RHS) of the \linkbeta.

A subterm is in \maybebeta{} if it has the shape \elpiIn{fapp[fuva
N | L]} and \elpiIn{distinct L} does not hold. In that case, \elpiIn{L} is split
in two sublist \elpiIn{PF} and \elpiIn{NPF} such that former is the longest
prefix of \elpiIn{L} such that \elpiIn{distinct PF} holds. \elpiIn{NPF} is the
list such that \elpiIn{append PF NPF L}. The LHS is set to a new variable
named $M$ with \elpiIn{PF} in scope whereas the RHS is given by the term
\elpiIn{app[uva N' PF | NPF]} where the \Ho variable identified by \elpiIn{N'}
is mapped to the \Fo{} variable named \elpiIn{N}.

\marginpar{signposting: a link is reconsidered in two cases. then make 2 paragraphs}
After its creation, a \linkbeta remain suspended until the head of the RHS is
instantiated by the oracle (see \cref{prop:simulation}). In this case the RHS is
$\beta$-reduced to a new term $t$. $t$ is either a term in \llambda, in
which case $t$ is unified with the LHS, otherwise, the link remain suspended
and no progress is performed. Another way to wake a \linkbeta up is when the
LHS is a term \elpiIn{T} and RHS has the shape \elpiIn{app[uva N PF | NPF]}
and some of the arguments in the \elpiIn{NPF} list become names. This is
possible after the resolution of other links. In this case, the list \elpiIn{L}
obtained by the concatenation between \elpiIn{PF} and \elpiIn{NPF} is split
again in to lists \elpiIn{PF'} and \elpiIn{NPF'}. If \elpiIn{PF} is not that
same as \elpiIn{PF'}, then we can 1) remove the current \linkbeta, 2) create a
new \linkbeta between \elpiIn{T} and \elpiIn{app[uva N' PF' | NPF']} and 3)
create a new \linketa between the variables \elpiIn{N} and \elpiIn{N'}.

\marginpar{A bit too fast, we first compile, then unify, then the oracle, then the manip...}
An example justifying this last link manipulation is given by the following
unification problem:

@@

We can now represent the \hrun execution from this configuration which will, at
first, dereference all the links, and then try to solve them. The only link
being modified is the second one, which is set to \textIn{X3 x =~$\beta$~= X2
x a}. The RHS of the link has now a variable which is partially in the PF,
we can therefore remove the original \linkbeta and replace it with the following 
couple on links:
